---
title: 'Assignment 10A: Sentiment Analysis'
author: "Jacob Shapiro"
date: "`r Sys.Date()`"
output: 
  html_document :
  df_print: paged
---
<style type="text/css">
.main-container {
  max-width: 100% !important;
  margin: auto;
}
options("width"=300)
</style>
<!-- Note need above for all columns to appear together -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r , include = FALSE}
library(dplyr)
library(tidyverse)
library(readr)
library(jsonlite)
library(rvest)
```

## Prompt

In Text Mining with R, Chapter 2 (https://www.tidytextmining.com/sentiment.html) looks at Sentiment Analysis.  In this assignment, you should start by getting the primary example code from chapter 2 working in an R Markdown document.  You should provide a citation to this base code.  You’re then asked to extend the code in two ways:

Work with a different corpus of your choosing, and
Incorporate at least one additional sentiment lexicon (possibly from another R package that you’ve found through research).

## Base code

```{r}
library(tidytext)

get_sentiments("afinn")

get_sentiments("bing")

get_sentiments("nrc")
```

```{r}
library(janeaustenr)
library(dplyr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)



nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

```{r}
library(tidyr)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

library(ggplot2)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
```

```{r}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

```{r}
get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)
```

```{r}
get_sentiments("bing") %>% 
  count(sentiment)
```


```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts

bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```

```{r}
custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```

```{r}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))


library(reshape2)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```

```{r}
p_and_p_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")

p_and_p_sentences$sentence[2]
```
```{r}
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())
```
```{r}
bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()
```

# Me adding

Using gutenbergr package. Load Edgar Allan Poe volume 2

```{r}
library(gutenbergr)

#see options to download works
gut_options <- gutenberg_works(
  languages = "en",
  only_text = TRUE,
  rights = c("Public domain in the USA.", "None"),
  distinct = TRUE,
  all_languages = FALSE,
  only_languages = TRUE
)

#chose Edgar Allan Poe - The Works of Edgar Allan Poe — Volume 2
poe <- gutenberg_download(2148)


#clean up blank rows
poe_clean<- poe[!(poe$text == ""),]

#label each line as line number
poe_clean <- poe_clean %>% mutate(linenumber = row_number())

#take out all punctuation and special characters
poe_clean$text <- gsub("[^[:alnum:] ]", "", poe_clean$text)

#imported in fragments - split on spaces to make list of words
# poe_clean %>%
#   mutate(split_values = str_split(text, "\\s+")) %>%
#   unnest_wider(split_values, names_sep = " ")
poe_clean <- poe_clean %>%   separate(text, into = c("word1", "word2", "word3", "word4", "word5","word6", "word7", "word8", "word9", "word10", "word11", "word12", "word13", "word14", "word15", "word16", "word17", "word18", "word19"), sep = "\\s+", extra = "merge", fill = "right")

#pivot into long list of words
poe_clean <- poe_clean %>% pivot_longer(cols = c("word1", "word2", "word3", "word4", "word5","word6", "word7", "word8", "word9", "word10", "word11", "word12", "word13", "word14", "word15", "word16", "word17", "word18", "word19"),
                      names_to = 'word_num',
                      values_to = "word")


#clean up blank and NA rows
poe_clean<- poe_clean[!(poe_clean$word == ""),]
poe_clean %>% na.omit()
```

## Sentiment Analysis

I can't figure out how to assign the different works so trying to just do it as a whole

```{r}
poe_sentiment <- poe_clean %>%
  inner_join(get_sentiments("bing")) %>%
  count(gutenberg_id, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

ggplot(poe_sentiment, aes(index, sentiment, fill = gutenberg_id)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~gutenberg_id, ncol = 2, scales = "free_x")
```
```{r}
afinn <- poe_clean %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  poe_clean %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  poe_clean %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

```{r}
bing_word_counts <- poe_clean %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts

bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```


