---
title: "Project 2 - Sachi - NYC Farmers Markets"
author: "Jacob Shapiro"
date: "`r Sys.Date()`"
output: 
  html_document :
  df_print: paged
---
<style type="text/css">
.main-container {
  max-width: 100% !important;
  margin: auto;
}
options("width"=300)
</style>
<!-- Note need above for all columns to appear together -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import wide text file

```{r , include = FALSE}
library(dplyr)
library(tidyverse)
library(readr)
```

Read in Sachi's NYC Farmers Markets data that is in csv file.

```{r}
farmers <- read_csv("https://raw.githubusercontent.com/jacshap/Data607/refs/heads/main/nyc_farmers_markets_visitors_wide.csv", col_names = TRUE, show_col_types = FALSE) 

knitr::kable(farmers)
```

## Format to long

```{r}
farmers <- farmers %>% pivot_longer(cols = 'Manhattan_Count':'StatenIsland_Count',
                               names_to = 'Borough',
                               values_to = "Count")
#clean up
farmers <- farmers %>% mutate(Borough = str_replace_all(Borough, c("_Count"="")))

knitr::kable(farmers)
```



## Analysis

Will do 2 analyses suggested by Sachi
1. Compare week-over-week trends by borough (levels + % change).
2. Compute share of total visitors each week (who’s contributing most).

### Week-over-week by borough

% change = (new value - old value)/old value * 100

```{r}
perc_change <- farmers %>% group_by(Borough) %>% mutate(lagged_perc = lag(Count)) %>% mutate(percent_change = (Count - lagged_perc)/lagged_perc*100) %>% select(Week, Borough, Count, percent_change)

perc_change <- perc_change[-(1:5),] %>% arrange(Borough)

knitr::kable(perc_change)

ggplot(data=perc_change, aes(x=Week, y=percent_change, fill = Borough)) + 
  geom_bar(stat="identity", position = "dodge") +
  labs(title = "Percent Change per Borough", x = "Week", y = "Percent Change") + 
  theme_light()
```

### Share of total visitors each week

```{r}
week_sum <- farmers %>% group_by(Week) %>% mutate(Sum_Week = sum(Count))

Visitor_Split_Percentage <- week_sum %>% mutate(Visitor_Split = Count/Sum_Week*100)

knitr::kable(Visitor_Split_Percentage)

ggplot(data=Visitor_Split_Percentage, aes(x=Week, y=Visitor_Split, fill = Borough)) + 
  geom_bar(stat="identity", position = "dodge") +
  labs(title = "Share of Visitors per Borough", x = "Week", y = "Percent of Visitors") + 
  theme_light()
```

